Deep Learning

### 2015-03-25 deepreview.tex
discussion about convexity

If ConvNets existed in the 1990s, why did they become so popular and
dominant only now? The idea of convexity of the objective function grew
as an attractive property of learning algorithms, thanks to the success
of kernel machines, which made neural nets (that involve a non-convex
training criterion) less appealing. As discussed below, it is only
in 2014 that theoretical and empirical evidence have suggested that local
minima are probably not as much of a problem as was initially thought~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-AISTATS2015}.
Much before that realization, though, a number of recent
ideas and technologies enabled the success. First, supervised ConvNets work best
when training data is abundant, and large labeled image datasets such
as ImageNet~\citep{imagenet_cvpr09} only appeared recently. Second,
the appearance of GPU cards capable of over 1Tflops, together with
GPU-optimized software for ConvNet made it possible to train these
large networks in a few weeks (now days) of computer time. Third, the
use of rectifying non-linearities instead of sigmoids allowed to train
much deeper networks than previously possible without unsupervised
pre-training~\cite{jarrett-iccv-09,Glorot+al-AI-2011-small}. Lastly,
the ``drop out'' regularization method~\citep{Srivastava14} was shown
to mitigate the effects of over-fitting in very large networks.

>>>>

If ConvNets and other neural nets existed in the 1990s, why did they become so popular and
dominant only now? It was initially thought that training of neural nets
would get stuck in poor solutions (called local minima) due to non-convexity
of the training objective, whereas the alternative learning framework
of {\em kernel machines}~\citep{Scholkopf+Smola-2002book} was more appealing
because it involved a convex optimization. As discussed below, it is only
in 2014 that theoretical and empirical evidence have suggested that local
minima are probably not as much of a problem as was initially 
thought~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-AISTATS2015}.
Much before that realization, though, a number of 
ideas and technologies over the last decade enabled the success. 
First, supervised ConvNets work best
when training data is abundant, and large labeled image datasets such
as ImageNet~\citep{imagenet_cvpr09} only appeared recently. Second,
the appearance of GPU cards capable of over 1Tflops, together with
GPU-optimized software for ConvNet made it possible to train these
large networks in a few weeks (now days) of computer time. Third, the
use of rectifying non-linearities instead of sigmoids allowed to train
much deeper networks than previously possible without unsupervised
pre-training~\cite{jarrett-iccv-09,Glorot+al-AI-2011-small}. Lastly,
the ``drop out'' regularization method~\citep{Srivastava14} was shown
to mitigate the effects of over-fitting in very large networks.
